Elaboration with First-Class Implicit Function Types

### 2020-03-01 paper.tex
minor fix

Curried functions are \emph{mostly} derivable from $\Nat$, $\top$ and $\Sigma$.
The type former is defined as follows:
\begin{alignat*}{3}
& \rlap{$\Pi^C : (A : \Tel) \to (\Rec\,A \to \U) \to \U$}\\
& \Pi^C (\zero,\,\_)\,B && :\equiv B\,\tt\\
& \Pi^C (\suc\,n,\,(A,\,B))\,C && :\equiv \{a : A\}\to \Pi^C\,(n,\,B\,a) (\lambda\,b.\,C\,(a,\,b))
\end{alignat*}
With this, we can also define $\mathsf{app} : \Pi^C\,A\,B \to (a : \Rec\,A) \to
B\, a$ and $\mathsf{lam} : ((a : \Rec\,A) \to B\, a) \to \Pi^C\,A\,B$, and all
equations in Figure \ref{fig:telescopes} hold definitionally, except
\LabTirName{curried-$\beta$} and \LabTirName{curried-$\eta$}. These do not hold
strictly, because $\Pi^C$, $\mathsf{app}$ and $\mathsf{lam}$ are all defined by
recursion on the $A$ telescope, but the $\beta\eta$ rules are specified
generically for an arbitrary (possibly neutral)
telescope. \LabTirName{curried-$\beta$} is still provable as a propositional
equality, and assuming function extensionality \LabTirName{curried-$\eta$} is
provable as well. For details, see our Agda formalization of these definitions,
which is included alongside the prototype implementation.

Hence, we can derive a somewhat weaker version of curried functions, with
propositional $\beta$ and $\eta$. From this, we still get consistency almost for
free. That is because whenever we have a model of the base theory with $\Nat$,
$\top$ and $\Sigma$, such that the model also validates equality reflection,
then every propositionally provable equation is also validated as a definitional
equation. And fortunately for us, standard models which prove consistency
usually validate equality reflection, e.g.\ set-theoretical models or standard
models in extensional type theory.

In contrast, showing canonicity, normalization and decidability of conversion
would require some extra work. We leave this to future work, but we expect that
it is straightforward to extend previous proofs to cover strict $\beta$ and
$\eta$ for curried functions.

>>>>

Curried functions are \emph{mostly} derivable from $\Nat$, $\top$ and $\Sigma$.
The type former is defined as follows:
\begin{alignat*}{3}
& \rlap{$\Pi^C : (A : \Tel) \to (\Rec\,A \to \U) \to \U$}\\
& \Pi^C (\zero,\,\_)\,B && :\equiv B\,\tt\\
& \Pi^C (\suc\,n,\,(A,\,B))\,C && :\equiv \{a : A\}\to \Pi^C\,(n,\,B\,a) (\lambda\,b.\,C\,(a,\,b))
\end{alignat*}
With this, we can also define $\mathsf{app} : \Pi^C\,A\,B \to (a : \Rec\,A) \to
B\, a$ and $\mathsf{lam} : ((a : \Rec\,A) \to B\, a) \to \Pi^C\,A\,B$, and all
equations in Figure \ref{fig:telescopes} hold definitionally, except
\LabTirName{curried-$\beta$} and \LabTirName{curried-$\eta$}. These do not hold
strictly, because $\Pi^C$, $\mathsf{app}$ and $\mathsf{lam}$ are all defined by
recursion on the $A$ telescope, but the $\beta\eta$ rules are specified
generically for arbitrary (possibly neutral)
telescopes. \LabTirName{curried-$\beta$} is still provable as a propositional
equality, and assuming function extensionality \LabTirName{curried-$\eta$} is
provable as well. For details, see our Agda formalization of these definitions,
which is included alongside the prototype implementation.

Hence, we can derive a somewhat weaker version of curried functions, with
propositional $\beta$ and $\eta$. From this, we still get consistency and
canonicity very cheaply. This is because models proving consistency or
canonicity for the base theory with $\Nat$, $\top$ and $\Sigma$ usually support
equality reflection. For consistency, standard set-theoretical models and models
in extensional type theory have this property, for canonicity, glued models
e.g.\ as in \cite{kaposi2019gluing, sterling2019algebraic} also have this property.

In contrast, showing normalization and decidability of conversion would require
some extra work. We leave this to future work, but we expect that it is not
difficult to extend previous proofs to cover strict $\beta$ and $\eta$ for
curried functions.
